---
title: "Lil_Rascals_Lab_12"
author: "Lil Rascals"
date: "April 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(cowplot)
library(stringr)
library(gapminder)

```
### Two Sample Permutation Test
Question: Is there a difference in life expectancy in Asian countries?
```{r}
gapminder3 <- gapminder

life_exp <- gapminder3 %>%
  select(lifeExp)
life_exp <- life_exp$lifeExp

perm_mean <- function(perms = 1000, value1, n1)
{
  ## Variables ##
  # perms: The number of permutations 
  # values (num): 
  # n1 (int): Size of group 1
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  mean_mat <- matrix(0,1000,1)
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly separate vector "values" into disjoint 
    # groups of size "n1" and "length(values) - n1" respectively
    rsamp <- sample(value1, n1, replace = FALSE)
    other <- sample(value1, length(value1)-n1, replace = FALSE)
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    A <- mean(rsamp)
    B <- mean(other)
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    mean_mat[[i]] <- (A-B)
  }
  
  # Step 5:
  # Return new updated vector, created in step 1
  return(mean_mat)
}
asia <- gapminder3 %>% filter(continent == "Asia") %>% select(lifeExp)

means <- as.tibble(perm_mean(value1 = life_exp, n1=200))
summary(asia)
summary(life_exp)
life_exp <- life_exp$lifeExp
realdiff <- mean(asia$lifeExp)-mean(life_exp)

std <- sqrt((sd(asia$lifeExp))^2/length(asia$lifeExp) + (sd(life_exp))^2/length(life_exp))
z_score <- realdiff/std
z_score
ggplot(means, aes(x=V1 ))+geom_histogram()+
  xlab("Means")+
  geom_vline(xintercept = realdiff, color = "red", size = .3)+
  ylab("Number")+
  ggtitle("Difference in Means in Life Expectancy")
```
Analysis: 
As you can see from the plot and the summary, the values of life expectancy are significantly higher for Asian countries on average than we would expect from various sample trials. Therefore, in Asian you have a higher life expectancy than in other places around the world. The mean of the random 1000 samples implies that there is no correlation between the difference in means of the random sample of the whole population, but the line shows that there is a difference in the real mean of Asian countries vs. other countries. 

The mean life expectancy around the world is 59.47 and the median is 60.71 whereas in Asia the mean is 60.06 and the median is 61.79 which is noticably higher. 
So we can conclude that the answer to our question is yes that life expectancy in Asian countries is much higher than the rest of the world.  

We also conducted a significance test and got a z-score of .8769146 which means a p-value of .19 so this result is significant at the .19 significance level. 
### Correlation Permutation Test
Question:
Is there a correlation between the GDP per Capita and life expectency?
```{r}

gapminder <- gapminder

gdp_percap <- gapminder%>%
  select(gdpPercap)

life_exp <- gapminder %>%
  select(lifeExp)

life_exp <- life_exp$lifeExp
gdp_percap <- gdp_percap$gdpPercap

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector 1 - for computing correlation
  # y: Vector 2 - for computing correlation
  ###############
  
  corr_mat <- matrix(0,1000,1)
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    rsamp <- sample(y, length(y), replace = FALSE)
    
    # Step 2:
    # Randomly mix up the values in the vector "y"
    
    corr_mat[[i]] = cor(x,rsamp)
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  return(corr_mat)
}
summary(gdp_percap)
summary(life_exp)
correlation <- as.tibble(perm_cor(x=gdp_percap,y=life_exp))
realcorr <- cor(gdp_percap, life_exp)
std <- sqrt((sd(gdp_percap))^2/length(gdp_percap) + (sd(life_exp))^2/length(life_exp))
z_score <- realcorr/std
ggplot(correlation, aes(x=V1))+geom_histogram()+
  geom_vline(xintercept = realcorr, color = "red", size = .3)+
  xlab("Correlation")+
  ylab("Number")+
  ggtitle("Correlation between GDP/Capita and Life Expectancy ")
```
Analysis:
The plot shows a normal distribution of data when tested over 1000 random samples. The mean of the sample data is centered around zero which would imply that the two factors are uncorrelated. This being said, some generated samples show a weak positive and a weak negative linear correlation. 
Overall, as you can see from the vertical line, there is a positive correlation between life expectancy and GDP per capita. This correlation is .5837062 which is a relatively strong positive correlation between the two variables implying that as one increases the other should also increase on average. 
We also calculated the z-score and got z=.00244 which means the p-value is about .50 which means these functions are correlated with a statistical significance at the .5
level. 

### Team Summary

Kelsey McKenna
I created the two tests in the function and made sure they worked correctly. I formulated the questions and created the analysis sections.
James Burke
I helped come up with the tests and how to do them. I also wrote part of the analysis.
Matthew Copeland
I helped with the questions we were trying to answer and made the graphs to complement the tests. 
Shreeya Basrur
I also helped with the questions and some of the analysis as well. 
